<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{rdpla use case: vizualize churches across DPLA holdings}
%\VignetteEncoding{UTF-8}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
	comment = "#>",
	collapse = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.path = "figure/"
)
```

# rdpla use case: vizualize churches across DPLA holdings

To aid users in what can be done with `rdpla`, the following is one example. 

Use case: 

> Search for all photos of churches and make a visualization through time
and facet by other factors

## Get item data

```{r}
library("rdpla")
library("data.table")
library("lubridate")
```

To get all data for `subject:churches` and of `type=image` we'd need to do about 100 
requests of size 500 records, but we'll just do one that so the example runs quickly. 

```{r}
res <- dpla_items(subject = "churches", type = "image", page_size = 500)
df <- res$data
```

## Parse data

```{r}
dates <- unlist(df$date, use.names = FALSE)
# remove 2nd part of a duration separted by " - ", just use first date
dates <- gsub("\\s-\\s.+", "", dates)
# remove 2nd part of a duration separted by "/", just use first date
dates <- gsub("/.+", "", dates)
# remove 2nd part of a duration separted by "-", just use first date
dates <- gsub("-[0-9]{4}", "", dates)
# remove any text, can't parse it
dates <- gsub("[[:alpha:]]+", "", dates)
# trim leading and trailing white space
dates <- gsub("^\\s+|\\s+$", "", dates)
# parse dates
df$date_corr <- parse_date_time(dates, c("Y", "ymd", "ym", "%Y-%m-%d%H:%M:%S%z"))
```

Make year column

```{r}
df$year <- year(df$date_corr)
```

Summarize by year

```{r}
df_dt <- as.data.table(df)
df_dt <- df_dt[order(year)]
df_sum <- df_dt[, list(count = length(id)), by = year]
df_sum <- as.data.frame(na.omit(df_sum))
```

## Vizualize

```{r tidy=FALSE, fig.width=8, fig.height=6}
library("ggplot2")

ggplot(df_sum, aes(x = year, y = count)) + 
  geom_bar(stat = "identity") + 
  theme_gray(base_size = 18)
```

Zooming in on the majority of the data

```{r tidy=FALSE, fig.width=8, fig.height=6}
ggplot(subset(df_sum, year > 1800), aes(x = year, y = count)) + 
  geom_bar(stat = "identity") + 
  theme_gray(base_size = 18)
```

Looks like there's a big spike in `r df_sum[which.max(df_sum$count), "year"]`

Let's have a look at a few of the titles

```{r}
sort(unique(subset(df, year == 1930)$title))[1:10]
```
