---
title: "Introduction to rdpla"
author: "Scott Chamberlain"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: TRUE
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Introduction to rdpla}
  \usepackage[utf8]{inputenc}
---

```{r echo=FALSE}
knitr::opts_chunk$set(
	comment = "#>",
	collapse = TRUE,
	warning = FALSE,
	message = FALSE
)
```

`rdpla`: R client for Digital Public Library of America

Metadata from the Digital Public Library of America ([DPLA](http://dp.la/)). They have [a great API](https://github.com/dpla/platform) with good documentation - a rare thing in this world. Further documentation on their API can be found on their [search fields](http://dp.la/info/developers/codex/responses/field-reference/) and [examples of queries](http://dp.la/info/developers/codex/requests/).  Metadata schema information [here](http://dp.la/info/wp-content/uploads/2013/04/DPLA-MAP-V3.1-2.pdf).

## Installation

Install from CRAN

```{r eval=FALSE}
install.packages("rdpla")
```

Development version

```{r eval=FALSE}
if (!requireNamespace("devtools")) {
  install.packages("devtools")
}
devtools::install_github("ropensci/rdpla")
```

Load `rdpla`

```{r}
library("rdpla")
```

## Search - items

> Note: limiting fields returned for readme brevity.

Basic search

```{r}
dpla_items(q="fruit", page_size=5, fields=c("provider","creator"))
```

Limit fields returned

```{r}
dpla_items(q="fruit", page_size = 10, fields=c("publisher","format"))
```

Limit records returned

```{r}
dpla_items(q="fruit", page_size=2, fields=c("provider","title"))
```

Search by date

```{r}
dpla_items(q="science", date_before=1900, page_size=10, fields=c("id","date"))
```

Search on specific fields

```{r}
dpla_items(description="obituaries", page_size=2, fields="description")
```

```{r}
dpla_items(subject="yodeling", page_size=2, fields="subject")
```

```{r}
dpla_items(provider="HathiTrust", page_size=2, fields="provider")
```

Spatial search, across all spatial fields

```{r}
dpla_items(sp='Boston', page_size=2, fields=c("id","provider"))
```

Spatial search, by states

```{r}
dpla_items(sp_state='Massachusetts OR Hawaii', page_size=2, fields=c("id","provider"))
```

Faceted search

```{r}
dpla_items(facets=c("sourceResource.spatial.state","sourceResource.spatial.country"),
      page_size=0, facet_size=5)
```

## Search - collections

Search for collections with the words _university of texas_

```{r}
dpla_collections(q="university of texas", page_size=2)
```

You can also search in the `title` and `description` fields

```{r}
dpla_collections(description="east")
```

## Visualize

Visualize metadata from the DPLA - histogram of number of records per state (includes __states__ outside the US)

```{r tidy=FALSE, fig.width=8, fig.height=6}
out <- dpla_items(facets="sourceResource.spatial.state", page_size=0, facet_size=25)
library("ggplot2")
library("scales")
ggplot(out$facets$sourceResource.spatial.state$data, aes(reorder(term, count), count)) +
  geom_bar(stat="identity") +
  coord_flip() +
  theme_grey(base_size = 16) +
  scale_y_continuous(labels = comma) +
  labs(x="State", y="Records")
```

